Aircraft Wildlife collisions

### 2021-07-16 aircraft_wildlife_doc.Rmd
Added color palette & proportional values

---
title: "Aircraft Wildlife collisions"
author: "Alexander Flick & Vy Nguyen"
date: "21/07/2021"
output:
  html_document: default
  pdf_document: default
---

@@

library(openintro)
data("birds")
library(ggplot2)
library(patchwork)
library(ggmosaic)

theme_set(theme_classic())
color_pal<-c("seagreen", "steelblue", "darkred", "orange")
theme.title<-theme(
    plot.title = element_text(
      hjust = 0.5, # center
      size = 12,
      color = "steelblue",
      face = "bold"
    ),
    plot.subtitle = element_text(
      hjust = 0.5, # center
      size = 10,
      color = "gray",
      face = "italic"
    )
  )

@@

ggplot(missing_df) + 
  aes(x=reorder(column, -count), y=count) +
  geom_bar(stat="identity") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

@@

```{r echo=FALSE, fig.asp=0.5}
birds_strike_phase <- ggplot(birds[!is.na(birds$phase_of_flt) & !is.na(birds$birds_struck),]) +
  aes(x = phase_of_flt, fill=birds_struck) +
  geom_bar()

#relationship of amount of birds struck with clouds?
birds_strike_sky <- ggplot(birds[!is.na(birds$sky) & !is.na(birds$birds_struck),])+
  aes(x=birds_struck, fill=sky)+
  geom_bar(position = "fill")

# Most collisions during the day #genauer betrachten
birds_strike_time <- ggplot(birds[!is.na(birds$time_of_day) & !is.na(birds$birds_struck),]) + 
  aes(x=time_of_day, fill=birds_struck) + 
  geom_bar()

### Birds struck out of birds seen
birds_struck_seen <- ggplot(birds[!is.na(birds$birds_seen) & !is.na(birds$birds_struck),]) +
  geom_mosaic(aes(x = product(birds_seen), fill=birds_struck)) +
  theme(axis.text.y=element_blank(), axis.ticks.y = element_blank())

@@

## Closer look on En Rout and Descent
```{r echo=FALSE, fig.asp=0.5, warning=FALSE}
above_ground <- birds[as.integer(birds$phase_of_flt)>3 & as.integer(birds$phase_of_flt)<8 & !is.na(birds$phase_of_flt),]
height_by_phase <- ggplot(above_ground) +
  aes(x = phase_of_flt, y = height, fill=phase_of_flt) +
  stat_boxplot(geom="errorbar", width=0.4) +
  geom_boxplot()
# Heights during Descent are on average higher than 
# hypothesis: Collisions happen more often during Descent and are therefore more prominent in the data set?

#plot height density for Descent and EnRoute
density <- ggplot(birds[birds$phase_of_flt== "Descent" | birds$phase_of_flt=="En Route",]) +
  aes(x = height, group=phase_of_flt,fill=phase_of_flt ) +
  geom_density(adjust=1.5,alpha=.8)

@@

## density birdsstuck by height
```{r echo=FALSE, warning=FALSE, fig.asp=0.5}
#could include only birds from vys categorization
#plot height density for Descent and EnRoute
density <- ggplot(above_ground[!is.na(above_ground$birds_struck),]) +
  aes(x = height, group=birds_struck,fill=birds_struck)+
  geom_density(adjust=1.5,alpha=.6) +
  xlim(0, 3500)
density

@@

```{r echo=FALSE, fig.asp=0.5}
#relationship of amount of birds seen with clouds?
birds_seen_plt <- ggplot(birds[!is.na(birds$time_of_day),])+
  aes(x=time_of_day, fill=birds_seen)+
  geom_bar(position = "fill") +
  theme(legend.position="none")

birds_seenNo_sky <- ggplot(birds[!is.na(birds$sky),])+
  aes(x=sky, fill=birds_seen)+
  geom_bar(position="fill")

@@

```{r echo=FALSE, fig.asp=0.5, figures-side, fig.show="hold"}
par(mfrow=c(1,2))
spineplot(birds$height, birds$birds_seen)
spineplot(birds$speed, birds$birds_seen)
```
The higher the speed & height, the less likely to see a bird before collision
for height > 2000 foot probability to see a birds before collision decreases 
for speed > 140 knots probability to see a bird before collision decreases

@@

# phase of flight by birdsstruck < birds seen
seen_vs_strike_phase <- ggplot(seen_vs_strike[!is.na(seen_vs_strike$phase_of_flt),]) +
  aes(x=phase_of_flt, fill=less_strike_than_seen) +
  geom_bar(position = "fill")

#ac_mass by leess-strike-than-seen (MOSAIC)
seen_vs_strike_ac_mass <- ggplot(seen_vs_strike[!is.na(seen_vs_strike$ac_mass),]) +
  geom_mosaic(aes(x=product(ac_mass), fill=less_strike_than_seen)) +
   theme(legend.position="none", axis.text.y=element_blank(), axis.ticks.y = element_blank())

#num_engs by leess-strike-than-seen (MOSAIC)
seen_vs_strike_num_engs <- ggplot(seen_vs_strike[!is.na(seen_vs_strike$num_engs),]) +
  geom_mosaic(aes(x=product(num_engs), fill=less_strike_than_seen)) +
   theme(legend.position="none", axis.text.y=element_blank(), axis.ticks.y = element_blank())

@@

```{r echo=FALSE, fig.asp=0.5}
#comparison right? because we actually want to compare here the positive impact
birds_struck_effect <- ggplot(birds[!is.na(birds$birds_struck) & !is.na(birds$effect),]) +
  aes(x = birds_struck, fill = effect) +
  geom_bar(position="fill")

@@

#mosaic plot better than barplot visualizes also absolute differences
birds_effect_acmass <- ggplot(birds[!is.na(birds$ac_mass) & !is.na(birds$effect),]) +
  geom_mosaic(aes(x = product(ac_mass), fill=effect)) +
  theme(legend.position="none", axis.text.y=element_blank(), axis.ticks.y = element_blank())

#birdds_seenYes by effect
birds_effect_seenYes <- ggplot(birds[!is.na(birds$effect),]) +
  geom_mosaic(aes(x = product(birds_seenYes), fill=effect))+
  #theme(legend.title = element_blank())
  theme(legend.position="none", axis.text.y=element_blank(), axis.ticks.y = element_blank())

>>>>

---
title: "Aircraft Wildlife collisions"
author: "Alexander Flick & Vy Nguyen"
date: "21/07/2021"
output:
  pdf_document: default
  html_document: default
---

@@

library(openintro)
data("birds")
library(ggplot2)
library(patchwork)
library(ggmosaic)
library("viridis")
library(RColorBrewer) # https://www.datanovia.com/en/blog/top-r-color-palettes-to-know-for-great-data-visualization/  -> Purples, BuPu; Map: BrBG

theme_set(theme_minimal())
color_pal<-c("seagreen", "steelblue", "darkred", "orange")
color_pal_sky<-c("lightgoldenrod","honeydew","skyblue3")
color_pal_TF<-c("darksalmon","cadetblue")
theme.title<-theme(
    plot.title = element_text(
      hjust = 0.5, # center
      size = 12,
      color = "steelblue",
      face = "bold"
    ),
    plot.subtitle = element_text(
      hjust = 0.5, # center
      size = 10,
      color = "gray",
      face = "italic"
    )
  )

@@

ggplot(missing_df) + 
  aes(x=reorder(column, -count), y=count) +
  geom_bar(stat="identity", fill="snow3") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

@@

```{r echo=FALSE, fig.asp=0.5}
birds_strike_phase <- ggplot(birds[!is.na(birds$phase_of_flt) & !is.na(birds$birds_struck),]) +
  aes(x = phase_of_flt, fill=birds_struck) +
  scale_fill_brewer(palette="Blues") +
  geom_bar()

#relationship of amount of birds struck with clouds?
birds_strike_sky <- ggplot(birds[!is.na(birds$sky) & !is.na(birds$birds_struck),])+
  aes(x=birds_struck, fill=sky)+
  scale_fill_manual(values = color_pal_sky)+
  geom_bar(position = "fill")

# Most collisions during the day #genauer betrachten
birds_strike_time <- ggplot(birds[!is.na(birds$time_of_day) & !is.na(birds$birds_struck),]) + 
  aes(x=time_of_day, fill=birds_struck) + 
  scale_fill_brewer(palette="Blues") +
  geom_bar()

### Birds struck out of birds seen
birds_struck_seen <- ggplot(birds[!is.na(birds$birds_seen) & !is.na(birds$birds_struck),]) +
  geom_mosaic(aes(x = product(birds_seen), fill=birds_struck)) +
  scale_fill_brewer(palette="Blues") +
  theme(axis.text.y=element_blank(), axis.ticks.y = element_blank())

@@

## Closer look on En Rout and Descent
```{r echo=FALSE, fig.asp=0.5, warning=FALSE}
above_ground <- birds[as.integer(birds$phase_of_flt)>3 & as.integer(birds$phase_of_flt)<8 & !is.na(birds$phase_of_flt),]
height_by_phase <- ggplot(above_ground) +
  aes(x = phase_of_flt, y = height, fill=phase_of_flt) +
  stat_boxplot(geom="errorbar", width=0.4) +
  scale_fill_brewer(palette="Pastel2") +
  geom_boxplot()
# Heights during Descent are on average higher than 
# hypothesis: Collisions happen more often during Descent and are therefore more prominent in the data set?

#plot height density for Descent and EnRoute
density <- ggplot(birds[birds$phase_of_flt== "Descent" | birds$phase_of_flt=="En Route",]) +
  aes(x = height, group=phase_of_flt,fill=phase_of_flt ) +
  scale_fill_brewer(palette='Pastel2') +
  geom_density(adjust=1.5,alpha=.8)

@@

## density birdsstuck by height
```{r echo=FALSE, warning=FALSE, fig.asp=0.5}
#could include only birds from vys categorization
#plot height density for Descent and EnRoute
density <- ggplot(above_ground[!is.na(above_ground$birds_struck),]) +
  aes(x = height, group=birds_struck,fill=birds_struck)+
  geom_density(adjust=1.5,alpha=.6) +
  scale_fill_brewer(palette="Blues") +
  xlim(0, 3500)
density

@@

```{r echo=FALSE, fig.asp=0.5}
#relationship of amount of birds seen with clouds?
birds_seen_plt <- ggplot(birds[!is.na(birds$time_of_day),])+
  aes(x=time_of_day, fill=birds_seen)+
  geom_bar(position = "fill") +
  scale_fill_brewer(palette="BuGn") +
  theme(legend.position="none")

birds_seenNo_sky <- ggplot(birds[!is.na(birds$sky),])+
  aes(x=sky, fill=birds_seen)+
  scale_fill_brewer(palette="BuGn") +
  geom_bar(position="fill")

@@

```{r echo=FALSE, fig.asp=0.5, figures-side, fig.show="hold"}
par(mfrow=c(1,2))
spineplot(birds$height, birds$birds_seen, col = rev(brewer.pal(n=3,name="BuGn")))
spineplot(birds$speed, birds$birds_seen, col = rev(brewer.pal(n=3,name="BuGn")))
```
The higher the speed & height, the less likely to see a bird before collision
for height > 2000 foot probability to see a birds before collision decreases 
for speed > 140 knots probability to see a bird before collision decreases

@@

# phase of flight by birdsstruck < birds seen
seen_vs_strike_phase <- ggplot(seen_vs_strike[!is.na(seen_vs_strike$phase_of_flt),]) +
  aes(x=phase_of_flt, fill=less_strike_than_seen) +
  scale_fill_manual(values = color_pal_TF)+
  geom_bar(position = "fill")

#ac_mass by leess-strike-than-seen (MOSAIC)
seen_vs_strike_ac_mass <- ggplot(seen_vs_strike[!is.na(seen_vs_strike$ac_mass),]) +
  geom_mosaic(aes(x=product(ac_mass), fill=less_strike_than_seen)) +
  scale_fill_manual(values = color_pal_TF)+
  theme(legend.position="none", axis.text.y=element_blank(), axis.ticks.y = element_blank())

#num_engs by leess-strike-than-seen (MOSAIC)
seen_vs_strike_num_engs <- ggplot(seen_vs_strike[!is.na(seen_vs_strike$num_engs),]) +
  geom_mosaic(aes(x=product(num_engs), fill=less_strike_than_seen)) +
  scale_fill_manual(values = color_pal_TF)+
  theme(legend.position="none", axis.text.y=element_blank(), axis.ticks.y = element_blank())

@@

```{r echo=FALSE, fig.asp=0.5}
#comparison right? because we actually want to compare here the positive impact
birds_struck_effect <- ggplot(birds[!is.na(birds$birds_struck) & !is.na(birds$effect),]) +
  aes(x = birds_struck, fill = effect) +
  scale_fill_brewer(palette="OrRd") +
  geom_bar(position="fill")

@@

#mosaic plot better than barplot visualizes also absolute differences
birds_effect_acmass <- ggplot(birds[!is.na(birds$ac_mass) & !is.na(birds$effect),]) +
  geom_mosaic(aes(x = product(ac_mass), fill=effect)) +
  scale_fill_brewer(palette="OrRd") +
  theme(legend.position="none", axis.text.y=element_blank(), axis.ticks.y = element_blank())

#birdds_seenYes by effect
birds_effect_seenYes <- ggplot(birds[!is.na(birds$effect),]) +
  geom_mosaic(aes(x = product(birds_seenYes), fill=effect))+
  scale_fill_brewer(palette="OrRd") +
  #theme(legend.title = element_blank())
  theme(legend.position="none", axis.text.y=element_blank(), axis.ticks.y = element_blank())
